---
title: "Project 2"
date: "2023-07-09"
author: "Spencer Williams & Stephen Macropoulos"
output: 
  github_document:
    toc: true
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The (Online News Popularity)[https://archive.ics.uci.edu/dataset/332/online+news+popularity] is a data set with a heterogeneous set of features about articles published by (Mashable)[www.mashable.com]. Multivariate data was gathered on sixty-one variables over a two year span. Our end result is to predict the number of shares in social networks. Below are some of the variables we will be looking at to help our prediction.

`Shares` - Number of shares (target)
`n_tokens_title` - Number of words in the title
`n_unique_tokens` - Rate of unique words in the content
`num_imgs` - Number of images
`num_videos` - Number of videos
`num_keywords` - Number of keywords in the metadata
`data_channel_is *` - There are six binary variables that will only be attributed to one variable. Theses include lifestyle, entertainment, business, social media, tech, and world.
`rate_positive_words` - Rate of positive words among non-neutral
`rate_negative_words` - Rate of negative words among non-neutral

# Libraries Used
** Leaving this blank for now. Will add as libraries are used **

# Reading in Data
```{r read, include=TRUE}

```